---
title: "BPC Bladder - Genomics / pathology"
author: "Alex Paynter"
date: "`r Sys.Date()`"
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---


```{r}
library(purrr)
library(here)
library(fs)
purrr::walk(.x = fs::dir_ls(here('R')), .f = source)

cpt_path_long <- readr::read_rds(
  here('data', 'cohort', 'cpt_path_long.rds')
)

cpt_path_sum <- cpt_path_long %>%
  group_by(cpt_genie_sample_id) %>%
  summarize(
    n_region = length(unique(map_custom)),
    regions = paste(sort(unique(map_custom)), collapse = ', '),
    .groups = 'drop'
  )

cpt_path_indicators <- cpt_path_long %>%
  select(cpt_genie_sample_id, map_custom) %>%
  mutate(is_there = T) %>%
  distinct(.) %>% 
  pivot_wider(
    names_from = map_custom,
    values_from = is_there,
    values_fill = FALSE
  ) %>%
  select(-`NA`)
```


```{r}

cpt_path_indicators %>% glimpse
# sort by colsums just to get good default printing:
cpt_path_indicators %<>%
  select(cpt_genie_sample_id, all_of(
    (cpt_path_indicators %>%
       select(-cpt_genie_sample_id)) %>%
       mutate(across(everything(), as.numeric)) %>%
       colSums(.) %>%
       sort(., decreasing = T) %>%
       names(.)
    )
  )
```

## Introduction

We were curious about genomic differences between metastatic sites.  First let's summarize the structure of PRISSMM as it relates to this question:

- Each cancer panel test (NGS test, `cpt_genie_sample_id`) is associated with one procedure and report pair (`path_proc_number`, `path_rep_number`), which we will call a path report.
- Each path report can note up to 30 sites that it is associated with.
- Sites in the path report are identified by ICD code, which we have grouped into custom body regions.

We'll start with an accounting of the distribution of body regions associated with each CPT report.

## CPT reports, sites

Of the `r nrow(cpt_path_sum)` NGS tests we have, `r nrow(filter(cpt_path_sum, n_region %in% 1))` have only one site/region included in the path sites.  Here is the list of regions for those with only one listed:

```{r}
#| include: true
filter(cpt_path_sum, n_region %in% 1) %>%
  group_by(regions) %>%
  summarize(n_region = sum(n_region)) %>%
  arrange(desc(n_region)) %>%
  flextable(.) %>%
  autofit(.)
```

Notes:

- Obviously a lot of these are "local" bladder samples and not mets.
- Of the specific non-bladder regions we're interested in (liver, bone, CNS, etc) we don't have a ton of samples to go on.

The following table lists the number of CPT tests associated with a particular site, regardless of whether it is the only site associated:

```{r}
#| include: true
cpt_path_indicators %>%
  select(-cpt_genie_sample_id) %>%
  tbl_summary(.)
```

## Genomics by site

Just to see if there's much here, we'll start by looking at the proportion of samples covering a pathology site which are DDR altered (oncogenic):

```{r}
# onco_alt_flags <- readr::read_rds(
#   here('data', 'genomic', 'ddr_def_compare', 'onco_alt_flags.rds')
# )

ddr_flags_first_sample <- readr::read_rds(
  here('data', 'genomic', 'ddr_def_compare', 'ddr_flags_first_sample.rds')
)

cpt_path_long <- left_join(
  ddr_flags_first_sample,
  cpt_path_long,
  by = c(sample_id = 'cpt_genie_sample_id')
)
```


```{r}
#| include: true
cpt_path_long %>%
  group_by(path_site = map_custom) %>%
  summarize(percent_ddr_alt = paste0(formatC(100 *mean(ddr), digits = 1, format = 'f'), '%')) %>%
  flextable(.) %>%
  autofit(.)
```

Notes:

- Possibly lower DDR alteration frequency in liver, lung, CNS.  Bladder samples were the highest.



