---
title: "BPC Bladder - DDR pathways"
author: "Alex Paynter"
date: "`r Sys.Date()`"
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---

```{r}
library(purrr); library(here); library(fs)
purrr::walk(.x = fs::dir_ls(here('R')), .f = source) # also loads lots of packages.
```

```{r}
dir_ddr <- here('data', 'genomic', 'ddr_def_compare')
gg_def_compare_heatmap <- readr::read_rds(here(dir_ddr, 'gg_def_compare_heatmap.rds'))
ft_ddr_def <- readr::read_rds(here(dir_ddr, 'ft_ddr_def.rds'))
gg_ddr_panel <- readr::read_rds(here(dir_ddr, 'gg_ddr_panel.rds'))

```

## Introduction

DDR pathway alterations have been hypothesized to have predictive value for those taking regimens containing both platinum-based chemotherapy and immunotherapy.  The definition of a gene being included in the DDR pathway varies widely depending on the institution, author and cancer type under study.  We will leverage both the urothelial carcinoma main GENIE subset (defined by oncotree code) and the subset selected for GENIE BPC to study the proportion who are "DDR altered" using various definitions.

### R4.1 Panel coverage

We (especially Michal) identified a few papers which have furnished DDR pathway gene lists.  These include:

- Teo 2017 (https://doi.org/10.1158/1078-0432.CCR-16-2520) - MSK
- Galsky 2018 (https://doi.org/10.1016/j.eururo.2017.12.001) - Mount Sinai.
- Iyer 2018 (http://ascopubs.org/doi/full/10.1200/JCO.2017.75.0158) - MSK.
- Yin 2020 (http://dx.doi.org/10.1634/theoncologist.2019-0851) - Ohio State.

Additionally, we created an empirical gene subset from GENIE.  This includes all genes that are a part of every GENIE BPC panel, and in at least one of the papers above.  We had to exclude small panels (those around 50 genes, used in few participants) to get a reasonable set, as several panels had 0 or 2 genes in the DDR pathway included.

The following heatmap shows the coverage of these four papers plus the intersection of GENIE panels (GENIE int):

```{r}
#| include: true
#| fig-height: 2
gg_def_compare_heatmap  
  

```

### R4.2 DDR-altered patient counts

Next we apply these gene lists to our data to find the number of people who have a DDR alteration.  We repeat this (in columns) for main GENIE and BPC, and using OncoKB to limit to only oncogenic alterations.

In addition to the definitions above we also include two more:

- Pearl, 2015 (https://doi.org/10.1038/nrc3891).  This is an academic ancestor of the papers above, and includes a huge number of genes which MAY be DDR-related.  This is included to give us a "maximal" comparison.  Two versions are included:
  - Pearl 2015 full:  every gene in the paper.
  - Pearl 2015 - select.  A subset of the pathways in the paper, specifically the pathways mentioned in the papers above ("MMR", "HR", "FA", "NHEJ", "BER", "NER").  The list of genes for each of these pathways is larger in the Pearl paper, so this is still a superset for comparison.

The Pearl lists were not included in the plot above mostly because they include too many genes and plot gets hard to read.

The following table shows the percentage having at least one ddr mutation (number of people):

```{r}
#| include: true
ft_ddr_def
```

**Notes**

- The main GENIE (GENIE-UC) and GENIE-BPC percentages are similar (but consistently slightly higher for BPC).  This gives some reassurance that we don't have an insanely biased sample.
- Comparing the "any alt." to "oncogenic" columns, the number of patients is generally 50-70% smaller when using the OncoKB oncogenic filter. 
- Except for the Pearl examples given for comparison, the proportion of main genie and BPC which has at least one oncogenic DDR alteration is between 19 and 24% for main genie and 18-26% for BPC.






```{r}
bpc_asco_2025_panel <- readr::read_rds(
  here('data', 'genomic', 'ddr_def_compare', 'bpc_asco_2025_panel.rds')
)
bpc_onco_mmr_flags <- readr::read_rds(
  here('data', 'genomic', 'ddr_def_compare', 'bpc_onco_mmr_flags.rds')
)
```


### R4.3 - Alteration frequency by gene

The following table shows the proportion of subjects whose first sample had an oncogenic alteration in each gene.  The genes included in this table are those in our 2025 ASCO manuscript definition of DDR altered.

```{r}
#| include: true
gtsummary::tbl_summary(
  select(bpc_asco_2025_panel, -c(sample_id, patient_id)) %>%
    select(sort(tidyselect::peek_vars()))
)
```

The following table does the same, for all the genes which were in the MMR pathway in some of the abstracts.

```{r}
#| include: true
gtsummary::tbl_summary(
  select(bpc_onco_mmr_flags, -c(sample_id, patient_id))
)
```

**Note:** Originally we included some of the genes which were in the Pearl paper but not those abstracts - the rates were basically zero for those so we removed that complication.  The above table does not consider the number of people tested for each of these genes.  The heatmap below shows that the common panels *do* cover these genes well though, so we're probably not drastically off as a result.









### R4.4 - Panel coverage for DDR genes

The following plot takes the genes included in the paper/definition comparison (R4.1) and shows the panel coverage in GENIE.  The axis is grouped to have the BPC-used panels at the top.  After that the axis is ordered by the number of samples tested (m).

The X axis puts the genes in the panel we used to define DDR first.  After that, it's alphabetical.  Coloring or similar is possible but ran out of time.

```{r}
#| include: true
#| fig-height: 8

gg_ddr_panel
```




```{r}
tmb_stack_ddr <- readr::read_rds(
  here(dir_ddr, 'tmb_stack_ddr.rds')
)
tmb_stack_mmr <- readr::read_rds(
  here(dir_ddr, 'tmb_stack_mmr.rds')
)
tmb_comp <- readr::read_rds(
  here(dir_ddr, 'tmb_comp.rds')
)

```


### R4.5 - TMB related to these panels

The goal of this section was to compare the tumor mutation burden of those with a DDR alteration to those who do not.  There's a problem with this - those with a DDR alteration have at least one mutation, while the cohort with no DDR alterations could have zero.  Therefore we're systematically biasing the outcome by having a deterministic link between number of mutations (TMB) and being DDR+.

A third column has been added to make a more fair comparison.  This is those without a DDR alteration, but who have at least one other mutation.  This way the control is bounded below by 1 mutation, just like the DDR+ positive group.  It would be my recommendation to use that column in our poster even if it is just a footnote or legend note.

The group of samples included is first samples for each index cancer in the BPC bladder cohort.

```{r}
#| include: true
tmb_stack_ddr %>% 
  select(grp, tmb_Mb, tmb_Mb_onco) %>%
  gtsummary::tbl_summary(
    ., 
    by = grp
  )
```
Here is the same table for MMR+: 

```{r}
#| include: true
tmb_stack_mmr %>% 
  select(grp, tmb_Mb, tmb_Mb_onco) %>%
  gtsummary::tbl_summary(
    ., 
    by = grp
  )
```

*Notes:*

- As expected the group with DDR and MMR alterations has higher TMB at all listed percentiles.
- TMB is listed in mutations per megabase (the usual).  The "onco" version only counts oncogenic mutations per megabase.  This is not standard really, but I think it is probably more relevant to our analyes.
- There are only 21 samples with zero mutations, so the issue I pointed out has a minor impact.



```{r}

```



