---
title: "BPC Bladder - Genomics / pathology"
author: "Alex Paynter"
date: "`r Sys.Date()`"
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---


```{r}
library(purrr)
library(here)
library(fs)
purrr::walk(.x = fs::dir_ls(here('R')), .f = source)

cpt_path_long <- readr::read_rds(
  here('data', 'cohort', 'cpt_path_long.rds')
)

cpt_path_sum <- cpt_path_long %>%
  group_by(cpt_genie_sample_id) %>%
  summarize(
    n_region = length(unique(map_custom)),
    regions = paste(sort(unique(map_custom)), collapse = ', '),
    .groups = 'drop'
  )
  

```

## Introduction

We were curious about genomic differences between metastatic sites.  First let's summarize the structure of PRISSMM as it relates to this question:

- Each cancer panel test (NGS test, `cpt_genie_sample_id`) is associated with one procedure and report pair (`path_proc_number`, `path_rep_number`), which we will call a path report.
- Each path report can note up to 30 sites that it is associated with.
- Sites in the path report are identified by ICD code, which we have grouped into custom body regions.

We'll start with an accounting of the distribution of body regions associated with each CPT report.

## CPT reports, sites

Of the `r nrow(cpt_path_sum)` NGS tests we have, `r nrow(filter(cpt_path_sum, n_regions %in% 1))` have only one site/region included in the path sites.  Here is the list of regions for those with only one listed:

```{r}
filter(cpt_path_sum, n_regions %in% 1) %>%
  dt_wrapper

```


```{r}

```




