---
title: "BPC Bladder - Genomic descriptives"
author: "Alex Paynter"
date: "Updated `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      warning = F, message = F,
                      fig.width = 7, fig.height = 5)
k_dpi <- 150
pal_pt_hc <- c("#004488","#ddaa33", "#bb5566")
```

<style type="text/css">
.main-container {
  max-width: 1000px !important;
  margin: auto;
}
</style>

```{r}
library(purrr)
library(here)
library(fs)

purrr::walk(.x = fs::dir_ls(here('R')), .f = source) # also loads lots of packages.
```



```{r set_gtsummary_theme}
theme_gtsummary_compact()
theme_gtsummary_language("en", big.mark = "") # for now.
```


```{r, load_datasets, warning = T}
read_wrap_clin <- function(p) {
  read_rds(file = here("data", 'cohort', p))
}

dft_pt <- read_wrap_clin("pt.rds")
dft_ca_ind <- read_wrap_clin("ca_ind.rds")
dft_cpt <- read_wrap_clin("cpt.rds")

read_wrap_geno <- function(p) {
  read_rds(file = here("data", 'genomic', p))
}

dft_gp_all <- read_wrap_geno('gene_panel_all.rds')
dft_gp_sum <- read_wrap_geno('gene_panel_sum.rds')
dft_gp_by_gene <- read_wrap_geno('gene_panel_by_gene.rds')
dft_onco_impact <- read_wrap_geno('oncokb_impact.rds')
```


## Content

```{r}
# Table explaining the sample type coding we'll use:
dfp_sample_type_expl <- dft_cpt %>% 
  group_by(sample_type) %>%
  summarize(
    num_sample = n(),
    # Each group should have one unique sample_type_simple_f - but if not this will fail loudly in the table. 
    coded_as = paste(unique(sample_type_simple_f)),
    .groups = "drop"
  ) %>%
  arrange(desc(num_sample))

# pp = per person.
dft_ngs_pp <- dft_cpt %>%
  group_by(record_id) %>%
  summarize(
    All = n(),
    Primary = sum(sample_type_simple_f %in% "Primary tumor"),
    Metastatic = sum(sample_type_simple_f %in% "Metastatic")
  ) 

dft_ngs_pp_long <- dft_ngs_pp %>%
  pivot_longer(
    cols = -record_id,
    names_to = "category",
    values_to = "count"
  ) %>%
  mutate(
    category = forcats::fct_inorder(category)
  )

gg_ngs_pp <- ggplot(
  data = dft_ngs_pp_long,
  aes(x = count, fill = category)
) + 
  geom_histogram(binwidth = 1, center = 0) +
  theme_classic() + 
  labs(x = "Number of tests",
       y = "Participants") + 
  scale_y_continuous(expand = expansion(add = 0, mult = c(0,0.05)),
                     n.breaks= 8) +
  scale_x_continuous(
    breaks = 0:10,
    expand = expansion(
      add = c(0,0), 
      mult = c(0,0.01)
    )
  ) + 
  facet_wrap(vars(category)) +
  theme(
    strip.text = element_text(hjust = 0),
    legend.position = "none"
  ) + 
  scale_fill_manual(
    values = pal_pt_hc
  )

n_prim_met_pair <- dft_ngs_pp %>% 
  filter(Primary >= 1 & Metastatic >= 1) %>%
  nrow(.)

n_gte_2_prim <- dft_ngs_pp %>% 
  filter(Primary >= 2) %>%
  nrow(.)

n_gte_2_met <- dft_ngs_pp %>% 
  filter(Metastatic >= 2) %>%
  nrow(.)
                            
```


### R2.1.1 NGS tests per person

This table uses the "Specimen Sample Type" variable from the Cancer Panel Test dataset.  We simplified this slightly (`coded_as`) to avoid small categories of tumors:

```{r, include = T}
dfp_sample_type_expl %>%
  huxtable(.) %>%
  huxtable::set_align(value = "left") %>%
  theme_compact(.)
```

**Discussion question:** Is there a reasonable way to group local recurrences with larger categories?

The following plot shows histograms of the number of tests per participant.  This is repeated for all tests (Primary, Metastatic, Other) and separately for Primary and Metastatic tumors.

```{r, include = T, fig.height = 3}
gg_ngs_pp
```

Observations:

- The overwhelming majority of participants have exactly one NGS test.
- About 400 people have no primary tumor tests at all.
- `r n_prim_met_pair` people have at least one primary and at least one metastatic tumor.
- Only `r n_gte_2_prim` participants have more than one primary tumor test.
- There are `r n_gte_2_met` with more than one metastatic test, overwhelmingly from MSK.






```{r}
# Add the number of patients and samples in parens:
dft_gp_all %<>%
  arrange(cpt_seq_assay_id) %>%
  left_join(
    ., 
    select(dft_gp_sum, cpt_seq_assay_id, n_pts, n_samples, n_genes),
    by = "cpt_seq_assay_id"
  ) %>%
  mutate(
    assay_lab = glue("{cpt_seq_assay_id} [{n_genes}] ({n_pts}, {n_samples})"),
    assay_lab = forcats::fct_inorder(assay_lab)
  ) 

# Get the list of genes covered in >50% of samples and which have at least 1 alteration in the data:

vec_gene_gte1_alt <- readr::read_rds(
  here('data', 'genomic', 'gene_gte1_alt.rds')
)
dft_gene_prop_samp_test <- readr::read_rds(
  here('data', 'genomic', 'gene_prop_samp_test.rds')
)
vec_test_prop_gte50 <- dft_gene_prop_samp_test %>%
  filter(prop >= 0.5) %>%
  pull(hugo)
dft_gp_all_limited <- dft_gp_all %>%
  filter(hugo %in% vec_gene_gte1_alt) %>%
  filter(hugo %in% vec_test_prop_gte50)


gg_panel_coverage <- plot_assay_heatmap(
  dat = dft_gp_all,
  plot_title = "Gene panel coverage"
)

gg_panel_coverage_limited <- plot_assay_heatmap(
  dat = dft_gp_all_limited,
  plot_title = "Gene panel coverage (limited)"
)
  
  
  
```




```{r}
gg_gene_sample_hist <- ggplot(
  dft_gp_by_gene,
  aes(x = prop_samp_tested)
) +
  geom_histogram(binwidth = 0.1, center = 0.05, fill = '#6699cc') +
  theme_classic() + 
  labs(x = "Proportion of samples gene tested in",
       y = "Number of genes") + 
  scale_y_continuous(expand = expansion(add = 0, mult = c(0,0.05)),
                     n.breaks= 8) +
  scale_x_continuous(
    labels = scales::percent,
    expand = expansion(add = c(0,0), mult = c(0,0.01)),
    breaks = seq(0,1, by = 0.1)) + 
  theme(
    # jump up the right margin to get the '100%' in 
    plot.margin = margin(unit(c(5.5, 10, 5.5, 5.5), "points"))
  )

# Quick computation for the observations below:
n_genes_total_coverage <- dft_gp_by_gene %>% filter(prop_samp_tested >= 0.999) %>% nrow()
n_genes_90_cov <- dft_gp_by_gene %>% filter(prop_samp_tested >= 0.899) %>% nrow()
n_genes_70_cov <- dft_gp_by_gene %>% filter(prop_samp_tested >= 0.699) %>% nrow()
dft_gp_by_gene %>% filter(prop_samp_tested < 0.5) %>% nrow()

```

### R2.1.2 Gene NGS coverage over samples

There are `r length(unique(dft_gp_by_gene$hugo))` genes covered by at least one panel in this dataset.  This is a histogram of genes, showing the number covering each decile of the proportion of samples (see example interpretation below).

```{r, include = T, fig.height = 4}
gg_gene_sample_hist
```

**Example interpretation:** The bar between 90% and 100% shows about 200 genes.  This means that approximately 200 genes were covered in at least 90% of panels.  Likewise, there were nearly 300 genes which were covered in under 10% of samples.

Observations:

- There are only `r n_genes_total_coverage` genes covered by every panel.  For our clinico-genomic analyses analyses it might be better to include genes covered by **most** panels to gather a larger set.  Two natural cut points based on the graph above may be:
  - Genes covered by at least 90% of panels: `r n_genes_90_cov`.
  - Genes covered by at least 70% of panels: `r n_genes_70_cov`.
  - Keep in mind that some of these will have too few positives to be used, further reducing the numbers of genes included.  We'll describe this in subsequent sections.
- The majority of genes are covered by under 50% of samples.  These will almost certainly need to be excluded for clinico-genomic analyses due to the degree of missingness.



### R2.1.3a Panel coverage

Heatmap of gene panel coverage.

```{r, include = T, fig.width = 10, fig.height = 5}
gg_panel_coverage
# Technically this works but it takes too long:
# plotly::ggplotly(gg_panel_coverage)
```

Observations:

- Panels vary in size considerably (under 50 to over 500). 







```{r}
n_pt <- dft_ca_ind %>% nrow
n_samp <- dft_cpt %>% nrow

gene_panel_table_help <- function(dat) {
  dat %>% 
    mutate(
      Patients = glue("{n_pts} ({round(n_pts/n_pt*100, 1)}%)"),
      Samples = glue("{n_samples} ({round(n_samples/n_samp*100, 1)}%)"),
      Genes = n_genes
    ) %>%
    select(
      Panel = cpt_seq_assay_id,
      Patients,
      Samples,
      Genes
    ) %>%
    distinct
}

dfp_gp <- gene_panel_table_help(dft_gp_all)
dfp_gp_limited <- gene_panel_table_help(dft_gp_all_limited)

gp_table_hux <- function(dat) {
  dat %>%
    huxtable(.) %>%
    huxtable::set_align(value = "left") %>%
    theme_compact(.)
}
  
```

**Table Version**  The following table shows similar information to the above figure: the number of patients, samples and genes covered by each panel.  

```{r, include = T}
gp_table_hux(dfp_gp)
```

*Note:* The patient percentages do not add to 100 (because each person can be tested with multiple panels).



```{r}
genes_panel <- dft_gp_all %>% 
  pull(hugo) %>% 
  unique %>%
  length


genes_panel_limited <- dft_gp_all_limited %>% 
  pull(hugo) %>% 
  unique %>%
  length
```


### R2.1.3b Panel coverage, limited

Same as R2.1.2a except we limit to genes with at least one alteration in the dataset, and genes that are covered in the panels used in at least 50% of samples.  This selection brings us down to `r genes_panel_limited` genes, from the original `r genes_panel`.

```{r, show_gene_table_limited, include = T}
gg_panel_coverage_limited
```

**Table version** skipped since most columns are identical to what was shown above.










```{r}
library(colorspace)

dft_onco_impact %<>% 
  mutate(n_div = if_else(
    oncogenic %in% c("Oncogenic", "Likely Oncogenic"),
    n,
    -n)
  ) %>%
  arrange(type, oncogenic)

# hcl_palettes("diverging", n = 5, plot = T)
pal_onco <- diverging_hcl(
  palette = "Tropic", n = 5
)




gg_onco_impact <- ggplot(
  dft_onco_impact,
  aes(x = n_div, fill = oncogenic, y = type)
) + 
  geom_col(
    position = "stack", orientation = 'y'
  ) +
  scale_fill_manual(
    values = pal_onco,
    name = NULL
  ) +
  guides(
    fill = guide_legend(reverse = T)
  ) + 
  labs(
    x = "Alterations (bulk count)",
    title = "Oncogenic annotation with OncoKB"
  ) + 
  theme_classic() + 
  theme(
    legend.position = "bottom",
    plot.title.position = "plot"
  )
```

### R2.2.1 Oncogenic annotation impact

We discussed annotating and potentially filtering the data using the "oncogenic" feature from oncoKB (https://www.oncokb.org/).  Looking at the alteration level (multiple alterations per sample are possible), the following plot shows the impact this would have on the raw data.

```{r, include = T, fig.height = 4}
gg_onco_impact
```

Observations:

- If we filtered our data to include only likely oncogenic or oncogenic alterations, we lose most of the mutations, and a smaller proportion or the fusions and CNAs.








## Upcoming

- Pathway analysis.
- Further charcterization of alterations after we decide on filtering:
  - Consequence (missense, frameshift, synonymous, etc).
  - Level of evidence (for likely oncogenic or oncogenic mutations).





## Problems/Concerns

### OncoKB Unknowns

In the prostate cohort we had some initial debate about whether it possibly correct that the majority of mutations have unknown oncogenic impact.  The answer to this was **Yes** from team members with a background at MSK (experienced with OncoKB), because at the end of the day OncoKB is people reading papers and inputting interesting things.  Negative (neutral impact) findings usually don't rate as interesting, so those don't end up in the data.